---
title: "Reproducible Research: Peer Assessment 1"
author: Rockgecko
date: 16-09-15
output: 
html_document:
keep_md: true
---

```{r setup, echo=FALSE, warning=FALSE}
## Set up
library(dplyr)
library(ggplot2)
data.archive<-"activity.zip"
data.filename<-"activity.csv"
options(scipen=4, digits=4)
```

## Loading and preprocessing the data
```{r load_data}
unzip("activity.zip",overwrite=TRUE)
data<-read.csv("activity.csv",header=TRUE)
```

## What is mean total number of steps taken per day?
First need to calculate total steps per day.
```{r sumbyday}
by.day<-group_by(data,date)
daily.data<-summarise(by.day,total.steps=sum(steps,na.rm=TRUE))

g<-ggplot(daily.data,aes(total.steps)) + 
    geom_histogram(col="red",alpha=0.2,binwidth=1000) +
    labs(title="Distribution of total steps per day") +
    labs(x="Number of steps",y="Count")
print(g)

mean.steps=mean(daily.data$total.steps,na.rm=TRUE)
median.steps=median(daily.data$total.steps,na.rm=TRUE)
```

The mean number of steps per day is `r mean.steps` and the median number of steps is `r median.steps`. 



## What is the average daily activity pattern?
```{r sumbyinterval}
by.interval<-group_by(data,interval)
interval.data<-summarise(by.interval,mean.steps=mean(steps,na.rm=TRUE))
g<-ggplot(interval.data,aes(x=interval,y=mean.steps)) + 
    geom_line(colour="red") +
    labs(title="Average steps per time interval") +
    labs(x="Interval",y="Average steps")
print(g)

max.interval=which.max(interval.data$mean.steps)
max.interval.starttime=interval.data$interval[max.interval]
```
The interval with the highest average number of steps is the interval labelled `r max.interval.starttime` (i.e. `r paste(max.interval.starttime,"<=time<",(max.interval.starttime+5))`).

## Imputing missing values
```{r missingvalues1}
num.NAs<-sum(is.na(data$steps))
```
There are `r num.NAs` missing values.
```{r missingvalues2}
NAs<-is.na(data$steps)
data$newsteps<-data$steps
data$newsteps[NAs]<-interval.data$mean.steps[NAs]

by.day<-group_by(data,date)
daily.data<-summarise(by.day,total.steps=sum(steps,na.rm=TRUE),total.newsteps=sum(newsteps,na.rm=FALSE))

g<-ggplot(daily.data,aes(total.newsteps)) + 
    geom_histogram(col="red",alpha=0.2,binwidth=1000) +
    labs(title="Distribution of total steps per day (missing data replaced)") +
    labs(x="Number of steps",y="Count")
print(g)

mean.newsteps=mean(daily.data$total.newsteps,na.rm=FALSE)
median.newsteps=median(daily.data$total.newsteps,na.rm=FALSE)
```
The mean number of steps per day is `r mean.newsteps` and the median number of steps is `r median.newsteps`.

SOMETHING WRONG - WHY STILL NAs?


## Are there differences in activity patterns between weekdays and weekends?
